Topological approach to finding blackholes in directed networks

### 2019-12-03 blackholes/pavt/article.tex
Algorithms description and pseudocode

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$CDG(C, E)$ -- channel dependency graph for a system topology with 
				direction order routing, \\
				min\_dir -- directions, which are less than this will not be considered}
			\Output{$CDG$ with added edges}
			// build the used direction sets\\ 
			\ForEach{$v \in C$}{
				$B(v) = \{s_i: \alpha_{u_i, s_i} \in BFS(CDG(C, E), v)\}$
			}
			\ForEach{$\alpha_{v_1, s_1}, \alpha_{v_2, s_2} \in C: v_2 = v_1 + s_1, (\alpha_{v_1, s_1}, \alpha_{v_2, s_2}) \notin E, s_1 \ge min\_dir$}{
				\If{$s_1 \notin B(\alpha_{v_2,s_2})$}{
					add $(\alpha_{v_1, s_1}, \alpha_{v_2, s_2})$ to $E$\\
					// rebuild used direction set for $\alpha_{v_1, s_1}$\\
					\ForEach{$v \in BFS(CDG(C, \bar{E}), \alpha_{v_1, s_1})$}{
						$B(v) = B(v) \cup B(\alpha_{v_2, s_2})$
					}
				}
			}
			\label{alg:add_additional_edge}
			\caption{Adding edges to CDG.}
		\end{algorithm}
	\end{center}
\end{figure}

@@

Let's consider two nodes in the closure of global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
special one should be topologically lower than non-special. Otherwise it would be one node basis.

To sum up: if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.

Imagine, we are given a graph and a set of nodes. What is the algorithm to decide whether these nodes form the basis of a blackhole? We shall decide if there's at least
one non-basic node.
First, we need to know what special nodes there are. To find out, we will take every global root of condensated graph and find topoological sort of its closure.
Those vertices, which suffice the rule above, we will mark as "special". 
This step can be precalculated.
Now, we shall calculate how many special nodes for every root there is in the given set. If there is more than one special node for root - the set is to be called non-basis and
omitted. Also we shall check if there are any non-special nodes lower than the special ones, and if those found - omit the node set.

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$G(V,E)$ -- directed acyclic graph (Graph condensation)}
			\Output{$Special$ -- }
			// build the used direction sets\\ 
			\ForEach{$v \in C$}{
				$B(v) = \{s_i: \alpha_{u_i, s_i} \in BFS(CDG(C, E), v)\}$
			}
			\ForEach{$\alpha_{v_1, s_1}, \alpha_{v_2, s_2} \in C: v_2 = v_1 + s_1, (\alpha_{v_1, s_1}, \alpha_{v_2, s_2}) \notin E, s_1 \ge min\_dir$}{
				\If{$s_1 \notin B(\alpha_{v_2,s_2})$}{
					add $(\alpha_{v_1, s_1}, \alpha_{v_2, s_2})$ to $E$\\
					// rebuild used direction set for $\alpha_{v_1, s_1}$\\
					\ForEach{$v \in BFS(CDG(C, \bar{E}), \alpha_{v_1, s_1})$}{
						$B(v) = B(v) \cup B(\alpha_{v_2, s_2})$
					}
				}
			}
			\label{alg:add_additional_edge}
			\caption{Adding edges to CDG.}
		\end{algorithm}
	\end{center}
\end{figure}

TopSortFilter algorithm:
\begin{enumerate}
    \item For every node 
    \item Calculate topological sort order
    \item Calculate closure size for every vertex
    \item For every node
\end{enumerate}
%
\section{Experimental Results}
To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey and TopSort) on a 
set of graphs.

>>>>

@@

Let's consider two nodes in the closure of global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
special one should be topologically lower than non-special. Otherwise it would be one node basis.
To sum up: if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.

Imagine, we are given a graph and a set of nodes. What is the algorithm to decide whether these nodes form the basis of a blackhole? We shall decide if there's at least
one non-basic node.
\begin{enumerate}
    \item First, we need to know what special nodes there are. This step can be precalculated. 
    To find out, we will take every global root of condensated graph and find topoological sort of its closure.
    Those vertices, which suffice the rule above, we will mark as "special". 
    \item Now, we shall calculate how many special nodes for every root there is in the given set. If there is more than one special node for root - the set is to be called non-basis and
    omitted. 
    \FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }
    \item Also we shall check if there are any non-special nodes lower than the special ones in related root topsort order, and if those found - omit the node set.
    When we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, omit this set.
    \FIXME{partially implemented: здесь условие выглядит несколько иначе: в упорядоченном наборе кандидатов специальная вершина может быть только первой. Впрочем, возможно, что тут баг. }
    \item Otherwise, we have a valid basis of a blackhole.
    To acquire blackhole we should, according to definition of basis, unite all the closures of basic nodes.
    Finally, we check for candidate set weak connectivity and if it is present - print blackhole.
\end{enumerate}

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$G(V,E)$ -- directed acyclic graph (Graph condensation) \\
                                $Cand(G)$ -- set of nodes; candidate to be basis}
			\Output{$True/False$}
                        // (1) precalc special nodes for every root \\
                        \ForEach{$r \in Roots(G)$} {
                            $Special_r = \emptyset$ \\
                            $tsOrder_r = TopSort(r)$ \\
                            \For{$i = 0; i<|tsOrder_r|; i=i + 1$} {
                                $v = tsOrder_r[i]$ \\
                                $C_v = Closure(v)$ \\
                                \If{$|C_v| == i$} {
                                    $Special_r = Special_r \cup v$ \\
                                }
                            }
                        }
			// (2) check for excluding specials \\
			\ForEach{$v \in Cand(G)$} {
                            \ForEach{$r \in Roots(G)$} {
                                \If{$v \in Special_r$} {
                                    $Spec_r = Spec_r \cup v$ \\
                                    \If{$|Spec_r| > 1$} {
                                        \Return $False$ \\
                                    }
                                }
                            }
			}
                        // (3) Check for candidates reachable from one another \\
                        \ForEach{$(v, u) \in Cand(G) \times Cand(G)$} {
                            \If{v is reachable from u} {
                                \Return $False$ \\
                            }
                        }
                        \Return $True$
			\label{alg:isbasis}
			\caption{BasisOrNot}
		\end{algorithm}
	\end{center}
\end{figure}

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$G(V,E)$ -- directed graph}
			\Output{$Blackhole$ -- set of blackholes of different sizes}

                        $Blackhole = \emptyset$\\
                        $Build\ graph\ condensation\ Cond(V',E')$\\
                        \For{$i = 1\ to\ |V'|$} {
                            \ForEach{$B_i \in V'(i)$} {
                                \If{$BasisOrNot(Cond(V',E'),B_i)$} {
                                    \If{$B_i$\ is\ weakly\ connected} {
                                        $BlackHole = Blackhole \cup B_i$ \\
                                    }
                                }
                            }
                        }
                        \Return $Blackhole$
			\label{alg:topsort}
			\caption{TopSort}
		\end{algorithm}
	\end{center}
\end{figure}

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$G$ -- directed graph \\
                                $V$ -- set of all nodes \\
                                $E$ -- set of all edges \\
                                $n$ -- max number of nodes each blackhole can contain}
			\Output{$Blackhole$ -- 1 to n-node blackhole set of G}

                        $Blackhole = \emptyset$ \\
                        \For{$i = 1\ to\ n$} {
                            $P_i = \{v | d_{out}(v) < i\}$ \\
                            \ForEach{$v \in P_i$} {
                                \If{$v \notin C_{i-1}$} {
                                    \If{$at\ least\ one\ of\ v's\ directed\ successors\ are\ not\ in\ P_i$} {
                                        $remove\ v\ from\ P_i$\\
                                        $remove\ all\ v's\ predecessors\ from\ P_i$\\
                                    }
                                }
                            }
                        }
                        $C_i = P_i$\\
                        \ForEach{$v \in C_i$} {
                            \If{$|Closure(v)| == i$} {
                                $Blackhole = Blackhole \cup Closure(v)$
                            }
                            \If{$|Closure(v)| >= i$} {
                                $remove\ v\ from\ C_i$ \\
                                $remove\ all\ v's\ predecessors\ from\ C_i$ \\
                            }
                        }
                        $F_i = C_i$\\
                        \ForEach{$B \in F_i(i)$} {
                            \If{$G(B)\ is\ weakly\ connected$} {
                                \If{$d_{out}(B) == 0$} {
                                    $Blackhole = Blackhole \cup B$ \\
                                }
                            }
                        }
                        \Return $Blackhole$
			\label{alg:iblackhole}
			\caption{iBlackHole}
		\end{algorithm}
	\end{center}
\end{figure}
%
\section{Experimental Results}
To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey \ref{alg:iblackhole} and TopSort ) on a 
set of graphs.
